Comparing tool variability and numerical variability in fMRI analyses

### 2021-09-27 paper/paper.tex
Add abstract and intro

@@

\begin{IEEEkeywords}
\end{IEEEkeywords}

@@

\begin{itemize}
    \item We presented a framework that can investigate the numerical instability of the pipelines based on Monte-Carlo arithmetic
    by creating a Fuzzy environment, so that instrument mathematical functions implemented in mathematical system libraries (libmath).

    \item Methodological choice can influence the final determining areas of brain activation.
    This opens so results flexibility~\cite{bowring2019exploring,bowring2021isolating}.

    \item In this paper, we aim to answer the questions 1) how the fMRI analyses are numerically stable?
    2) how the numerical variability is in comparison with the tool variability?
\end{itemize}

@@

Finally, the instrumented functions are loaded in the pipeline using LD\_PRELOAD, a Linux environment variable
to force load a shared library into an executable. ​
This allows functions defined in FL to transparently
overload the original ones without the need to modify or recompile the analysis pipeline.​

@@

@@

@@

% Fig2
In Figure~\ref{fig:dice-thresh}, we compare region-by-region Dice coefficients of
the group-level thresholded maps for all pairs of software packages in BT and WT.
This shows a linear correlation between Dice values in BT and WT,
which implies that the variability of activated voxels in both conditions is similar.
Also, small P-values across all three pairs, including \num{1e-10} for \fslafni, \num{6e-04} for \fslspm,
and \num{2e-05} for \afnispm, confirms this correlation of similarities.
The vertical line where the Dice score is zero in BT shows regions in the brain with no overlaps between activated voxels in BT but WT,
including left/right `V1\_ROI' in \fslafni and \fslspm, and left/right `TGd\_ROI' in \afnispm results.

@@

\subsection{Group-level unthresholded maps}

@@

% Fig4
The scatter plot in Figure~\ref{fig:correlations} represents the correlation of standard deviation in WT and BT variability.
We found two major clusters, including the identity cluster that corresponds to the correlations
between BT and WT with the ratio of $0.5 < BT/WT < 1.5$,
and the upper cluster that shows voxels where BT $\approx$ 0.
The percentage of voxels included in the identity cluster is \%9.9 in \fslspm, \%17.3 in \fslafni, and \%13.8 in \afnispm,
and total voxels in the upper cluster are $\approx$ \%1.
The identity area is also represented on the MNI space in the second row in Figure~\ref{fig:correlations}.
This figure shows the spatial localization of the parts of the brain that BT variability is correlated with the numerical variability.
This refines the presented results in Figure~\ref{fig:unthresh-varmaps},
which indicates how correlation is uniformly distributed across the brain.

@@

>>>>

Computational environments are known to affect the pipeline results in neuroimaging,
presumably due to the creation, propagation, and amplification of small numerical errors.
Numerical instability is a characteristic of the pipelines which amplify numerical errors
originating in floating-point computations and then hamper reproducibility of the analyses.
% It has been widely investigated that methodological flexibility can influence the final determined areas of brain activation.
In particular, errors that highlight numerical instabilities are likely to appear as a result of changes in data analysis software packages.
This work presents a comparative assessment of the group-level fMRI analysis between tool variability and numerical variability.   
We characterize the effect of numerical perturbations on the fMRI analysis using the Fuzzy Libmath (FL) framework.
FL is a technique to instrument the mathematical functions embedded in mathematical libraries based on the Monte-Carlo Arithmetic (MCA).
We show that by taking advantage of the Fuzzy libmath framework, one can evaluate the numerical uncertainty of the pipelines. 
We (1) evaluated the numerical stability of the fMRI analysis that has been reproduced using three of the most common tools in neuroimaging, FSL, AFNI, and SPM,
and (2) compared between tool variability and numerical variability.
We found a significant numerical variability on both thresholded and unthresholded images;
results show that the order of magnitude of variations between tools was highly more than numerical variability.
Also, we found a linear correlation between Dice scores computed region-by-region on thresholded maps in both conditions.

@@

\begin{IEEEkeywords}
  Numerical Instability, Reproducibility, Monte-Carlo Arithmetic, Neuroimaging
\end{IEEEkeywords}

@@

% Data analysis workflows in many scientific domains have become increasingly complex and flexible. 
Recent studies highlighted the instability of the neuroimaging pipelines depending on the computing platform,
software package, and even tool versions. Changes in the computational
environment such as compilers, libraries, operating systems may introduce small numerical errors and create
significantly different results in unstable pipelines~\cite{Glatard2015,Gronenschild2012,salari2020spot}.
Moreover, the impact of methodological changes on fMRI analyses has been investigated extensively~\cite{bowring2019exploring,botvinik2020variability,bhagwat2021understanding,carp2012plurality}.
For instance, in related works, it has been shown that running the same fMRI experiments by different teams can substantially affect
scientific conclusions~\cite{botvinik2020variability,carp2012plurality};
replication of fMRI experiments using the three most well-known software packages can influence the final determining areas of
brain activations~\cite{bowring2019exploring};%bowring2021isolating. 
also, the choice of preprocessing pipelines on neuroimaging cortical surface analyses is compound with the instabilities~\cite{bhagwat2021understanding}.
% In the presence of such instabilities, it is often hard to trust the data processing results. % validity of the computational results.

In such a heterogeneous environment, numerical instability is an essential issue for reproducibility.
Numerical instability is a characteristic of pipelines that results from the influence of the floating-point arithmetics
and iterative convergence of numerical errors~\cite{freitas2002issue}.
Stochastic arithmetic approaches such as Monte-Carlo Arithmetic (MCA)~\cite{Parker1997-qq} have been used to study the impact of numerical errors
originating in floating-point computations in mathematical libraries.
In~\cite{salari2021accurate}, we quantified the numerical stability of the HCP preprocessing pipeline~\cite{glasser2013} based on the MCA method by creating a Fuzzy environment
, so that instrument mathematical functions are implemented in mathematical system libraries.
As a result of numerical perturbations, we discovered a very low numerical precision in the result images comparable to the operating system variability.
In a related study~\cite{kiar2020numerical}, the instability of results was explored by instrumenting a connectome estimation pipeline with the MCA technique.
These works demonstrate the necessity of numerical uncertainty quantification for understanding related issues that hamper the computational reproducibility of analyses.

In this work, we reproduce an fMRI experiment with different neuroimaging software packages in the presence of the Fuzzy environment,
and then quantify the numerical variability and between tool variability in the results. 
The primary objective of this study is to answer these two questions: 1) how the fMRI analyses across tools are numerically stable?
2) how the numerical variability is in comparison with the tool variability?
This comparative study reveals the importance of numerical variability and motivates research studies to evaluate the numerical uncertainty of the pipelines.

% We start to reproduce an analysis from Bowring, this can be as a practice for reproducibility manner in the community.
% In particular, we investigate the effect of 1) between software 2) within software (numerical)
% We present a comparative assessment of group-level analysis of an fMRI pipeline.

@@

Finally, the instrumented functions are loaded in the pipeline using LD\_PRELOAD, a Linux environment variable
to force load a shared library into an executable.
This allows functions defined in FL to transparently
overload the original ones without the need to modify or recompile the analysis pipeline.

@@

% MCA allows for three perturbation modes: Random Rounding (RR) introduces the
% perturbation in function outputs, simulating roundoff errors; Precision Bounding
% (PB) introduces the perturbation in function operands, allowing for the
% detection of catastrophic cancellations; and, Full MCA combines RR and PB,
% resulting in the following perturbation:
% \begin{equation} \label{eq:mca_modes}
%   mca\_mode(x \circ y) = inexact_{RR}(  inexact_{PB}(x) \circ inexact_{PB}(y) )
% \end{equation}

@@

% In this case, differences in the number of threads raised floating point rounding errors.
% So it's probabilistic to have this issue by changing the hardware architecture like variation in CPU, GPU, APU from intel, AMD, or Nvidia.
% It still may suffer from Hyper-threading (often on intel CPUs), means a physical CPU core divides into different
% virtual cores that allows more than one thread to run on each core. It says that you allocated one CPU, but virtually have multi-threads.

% EC is a means to assess whether only superficial scaling differences (differences by a scale factor over all voxels) 
% are responsible for disparities between pair of images.

@@

% Fig2
In Figure~\ref{fig:dice-thresh}, we compare region-by-region Dice coefficients of
the group-level thresholded maps for all pairs of software packages in BT and WT.
This shows a linear correlation between Dice values in BT and WT,
which implies that the variability of activated voxels in both conditions is similar.
Also, small P-values across all three pairs, including \num{1e-10} for \fslafni, \num{6e-04} for \fslspm,
and \num{2e-05} for \afnispm, confirms this correlation of similarities.
The vertical line where the Dice score is zero in BT shows regions in the brain with no overlaps between activated voxels in BT but WT,
including left/right `V1\_ROI' in \fslafni and \fslspm, and left/right `TGd\_ROI' in \afnispm results.
% The regression line is computed over Dices where BT and WT are not zero.
% The average of Dice scores range from to 
% FSL-AFNI variations has the most similarity correlations which means that the results of these two tools are more \dots

% In response to the out of border activated voxels can say:
% This is likely due to the fact that SPM consistently had the smallest analysis mask out of the three packages,
% while FSL had the largest. Number of activated voxels prove this fact.

@@

\subsection{Group-level unthresholded maps}
% \subsection{Variability of unthresholded statistical maps}

@@

% Fig4
The scatter plot in Figure~\ref{fig:correlations} represents the correlation of standard deviation in WT and BT variability.
We found two major clusters, including the identity cluster that corresponds to the correlations
between BT and WT with the ratio of $0.5 < BT/WT < 1.5$,
and the upper cluster that shows voxels where BT $\approx$ 0.
The percentage of voxels included in the identity cluster is \%9.9 in \fslspm, \%17.3 in \fslafni, and \%13.8 in \afnispm,
and total voxels in the upper cluster are $\approx$ \%1.
The identity area is also represented on the MNI space in the second row in Figure~\ref{fig:correlations}.
This figure shows the spatial localization of the parts of the brain that BT variability is correlated with the numerical variability.
This refines the presented results in Figure~\ref{fig:unthresh-varmaps},
which indicates how correlation is uniformly distributed across the brain.
% Spatial localization of significant activation in the thresholded images also varied across software packages.
% Moreover, results show that pair of FSL-SPM has the least correlation of variability in within tool and between tool.
% This will implies that ... which can help for further investigations.

@@

% There are many statistical comparisons, but the neuro-scientific interpretation of results is not on my side.
